---
title: "Assignment 3 - Fundamentals of Machine Learning"
author: "Julia Thacker"
date: "10/12/2021"
output: html_document
---

```{r}

require(e1071)
require(caret)
UniversalBank<-read.csv("UniversalBank.csv")
```
Loaded packages needed and read the data
```{r}
UniversalBank$CreditCard<-factor(UniversalBank$CreditCard)
UniversalBank$Online<-factor(UniversalBank$Online)
UniversalBank$Personal.Loan<-factor(UniversalBank$Personal.Loan)
```
Converted each of the necessary variables into factors
```{r}
selected.var<-c(10,13,14)
set.seed(80)
bank.train.index = createDataPartition(UniversalBank$Personal.Loan,p=0.6, list = FALSE)
bank.train = UniversalBank[bank.train.index,selected.var]
bank.valid<-UniversalBank[-bank.train.index,selected.var]
summary(bank.train)
summary(bank.valid)
```
Selected the required variables, set seed, and partitioned the data into 60% training and 40% validation
```{r}
attach(bank.train)
ftable(CreditCard,Personal.Loan,Online)
```

Created a pivot table of the training data with Online as a column variable and Credit Card and Personal Loan as row variables.

The probability that a customer who owns a bank credit card and is actively using online banking services would accept the loan offer is 45/513 or 8.77%. This is because there is a total of 513 customers in this data set that both own a bank credit card and are actively using online banking services. Out of those 513 customers, only 45 accepted the loan.
```{r}
ftable(Personal.Loan,Online)
ftable(Personal.Loan,CreditCard)
```
Created two pivot tables of the training data - one with Personal Loan as the rows and Online as the columns, and one with Personal Loan as the rows and Credit Card as the columns. 
```{r}
prop.table(table(bank.train$Personal.Loan,bank.train$Online),margin=1)
prop.table(table(bank.train$Personal.Loan,bank.train$CreditCard),margin=1)
```
Also made them as pivot tables that generate proportions instead of count to be able to evaluate their probabilities.

```{r}
83/288
175/288
288/3000
803/2712
1605/2712
2712/3000
```
Computed:
i.P(CC = 1 | Loan = 1) = 28.8%
ii.P(Online = 1 | Loan = 1) = 60.8%
iii.P(Loan = 1) = 9.6%
iv.P(CC = 1 | Loan = 0) = 29.6%
v.P(Online = 1 | Loan = 0) = 59.2%
vi.P(Loan = 0) = 90.4%
```{r}
(0.2881944*0.6076389*0.096)/((0.2881944*0.6076389*0.096)+(0.2960914*0.5918142*0.904))
```
8.7% in the pivot table in question B compared to 9.6% in the naive bayes computation. These two probabilities are similar, but the pivot table answer would be a more accurate estimate because the pivot table is calculating an exact answer based on all of the data, but the naive bayes is using assumptions I have made.

```{r}
bank.nb<-naiveBayes(Personal.Loan ~ .,data = bank.train)
bank.nb
```
Computed the Naive Bayes probability on the training data.
The probability of the loan being accepted is 9.6% and the probability that the loan is not accepted is 90.4%. This 9.6% is the same value that I calculated manually in the previous method.
