---
title: "Final Exam - Fundamentals of Machine Learning"
author: "Julia Thacker"
date: "12/01/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(e1071)
airlinedata<-read.csv("airline_passenger_satisfaction.csv")
airlinedata<-na.omit(airlinedata)
```
Read the data and removed any NAs
```{r}
airlinedata$Gender<-as.factor(airlinedata$Gender)
airlinedata$type_of_travel<-as.factor(airlinedata$type_of_travel)
airlinedata$customer_class<-as.factor(airlinedata$customer_class)
airlinedata$customer_type<-as.factor(airlinedata$customer_type)
airlinedata$satisfaction<-as.factor(airlinedata$satisfaction)
airlinedata$inflight_wifi_service<-as.factor(airlinedata$inflight_wifi_service)
airlinedata$inflight_entertainment<-as.factor(airlinedata$inflight_entertainment)
```
Converted necessary variables into factors
```{r}
 str(airlinedata)
```

```{r}
library(ggplot2)
ggplot(airlinedata,aes(age, colour =satisfaction))+geom_freqpoly(binwidth=1)
```
Plotted the count of passengers that were neutral or dissatisfied vs the ampunt of satisfied passengers by age.
Younger people were more likely to be dissatisfied.
```{r}
set.seed(123)
airline.train.index = createDataPartition(y=airlinedata[,1],p=0.5)[[1]]
airline.train = airlinedata[airline.train.index,]
airline.valid<-airlinedata[-airline.train.index,]
summary(airline.train)
summary(airline.valid)
```
Partitioned the data into 50% training data and 50% validation data.
```{r}
ftable(airline.train$satisfaction,airline.train$inflight_entertainment)
```
Created a pivot table of the inflight entertainment survey results based on customer satisfaction.

```{r}
ftable(airline.train$satisfaction,airline.train$inflight_wifi_service)
```
Created a pivot table of the in flight WiFi service survey results based on customer satisfaction.

People that did not use the WiFi service were primarily still satisfied overall.
More passengers that rated the WiFi service poorly ended up being dissatisfied overall.
There is a very small amount of people that rated the WiFi service a 5 and ended up being dissatisfied overall.

```{r}
round(prop.table(table(airline.train$satisfaction,airline.train$inflight_entertainment),margin=1),2)
round(prop.table(table(airline.train$satisfaction,airline.train$inflight_wifi_service),margin=1),2)
```
Created a pivot table for each service that shows the probabilities.

A lower rating given for in flight WiFi service was more likely to result in overall dissatisfaction.

```{r}
variables<-c(8,15,24)
set.seed(123)
airline.train.index2 = createDataPartition(airlinedata$satisfaction,p=0.5,list=FALSE)
airline.train2 = airlinedata[airline.train.index2,variables]
airline.valid2<-airlinedata[-airline.train.index2,variables]
```
Partitioned the data again using only the 3 necessary variables.
```{r}
round(prop.table(table(airline.train2$satisfaction,airline.train2$inflight_entertainment),margin=1),2)
round(prop.table(table(airline.train2$satisfaction,airline.train2$inflight_wifi_service),margin=1),2)
```
Created two pivot tables of this data partition to show the probabilities of each outcome.

```{r}
airline.nb<-naiveBayes(airline.train2$satisfaction ~ .,data = airline.train2)
airline.nb
```
Computed the Naive Bayes probability.

The probability that a customer will be neutral or dissatisfied is 57%. The probability that a customer will be satisfied is 43%.

